---
title: "Lab 4 -- Visualizing Model Assumptions"
author: "Your Group Names Here (include last names!)"
date: "Due: September 24 by 5pm"
output: html_document
---
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)

drunk <- read_xlsx("data/Impaired_Driving_Death_Rate.xlsx")

```

For this lab we will look at the rate of deaths by age/gender 
(per 100,000 population) for people killed in crashes involving a driver 
with BAC =>0.08%. These data come from the CDC (Center for Disease Control and 
Prevention) for the years 2012 and 2014.  

Data for 2012 were obtained from the Fatality Analysis Reporting System (FARS) 
and data for 2014 were obtained from the National Highway Traffic 
Administration's (NHTSA) Fatality Analysis Reporting System (FARS). Blank cells 
indicate data are suppressed. Fatality rates based on fewer than 20 deaths are suppressed.


# Modeling Differences in Fatalities between 2012 and 2014   

We could investigate if the rate of deaths (per 100,000 people)  
from drunk driving has changed between 2012 and 2014.   

```{r}

drunk_all_years <- drunk %>% 
  # selecing only the two columns with data on all deaths for 2012 & 2014 
  select(`All Ages, 2012`, `All Ages, 2014`) %>% 
  # gathering the two columns into one column called year and another called deaths  
  gather(`All Ages, 2012`:`All Ages, 2014`, key = "year", value = "deaths") %>% 
  # changing the year column to be only 2012 and 2014 (removing the All Ages prefix)
  mutate(year = case_when(year == "All Ages, 2012" ~ "2012", 
                          year == "All Ages, 2014" ~ "2014")) 
  
```  

Alright, to investigate the differences in the fatality rates in 2012 versus 
2014, we could conduct a permutation test or a t-test.  

__1. What are the assumptions of a permutation test?__  
</br>



</br>
</br>

__2. What assumption does a t-test have that a permutation test does not?__ 
</br>



</br>
</br>


## Visualizing Model Assumptions  

```{r, warning = FALSE}
drunk_all_years %>% 
  # pipe dataset into the plot
  ggplot(aes(x = year, y = deaths)) + 
  # specify x and y variables 
  geom_violin(aes(fill = year)) + 
  # specify how the violins should be colored 
  geom_jitter(width = 0.1) + 
  # adds data points on top of the violins   
  theme(legend.position = "none")  
  # removes legend from plot, since it is redundant with x-axis label
```

__3. What assumption does this plot help you assess?__ 
</br>



</br>
</br>


The assumption of normally distributed residuals (or the responses) 
can be kind of tricky to figure out! Let's look at the same plot as above, 
just flipped around. This coordinate flip makes it a bit easier to visualize 
the distributions of the fatality rates for each year.    

```{r, warning = FALSE}
drunk_all_years %>% 
  # pipe dataset into the plot
  ggplot(aes(x = year, y = deaths)) + 
  # specify x and y variables 
  geom_violin(aes(fill = year)) + 
  # specify how the violins should be colored 
  geom_jitter(width = 0.1) + 
  # adds data points on top of the violins   
  coord_flip() + 
  # flips the coordinates (x to y, y to x)
  theme(legend.position = "none")  
  # removes legend from plot, since it is redundant with x-axis label
```

__4. Do you believe the assumption of normally distributed responses is violated? Why or why not?__ 
</br>


</br>
</br>


## Training Models 


```{r, warning = FALSE}
library(infer)

# CONDUCTING A T-TEST

drunk_all_years %>% 
  # t_test gives a tidier output than t.test
  t_test(deaths ~ year, # specify response ~ explanatory 
         order = c("2014", "2012"),  # specify order of subtraction
         alternative = "two_sided",  # specify direction of alternative 
         mu = 0, # specify mean in null hypothesis (if 2104 = 2012, then 2014 - 2012 = 0)
         conf_int = TRUE, # specify if you want a confidence interval (TRUE = yes, FALSE = no)
         conf_level = 0.99 #specify the confidence level to be used for the interval 
  )


# CONDUCTING A PERMUATION TEST  

# Step 1: Calculate observed difference in mean fatality 
obs_diff <- drunk_all_years %>% 
  specify(deaths ~ year) %>% 
  calculate(stat = "diff in means", 
            order = c("2014", "2012")
            )

# Step 2: Generate permuted differences in mean fatality  
permute <- drunk_all_years %>% 
  specify(deaths ~ year) %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 5000, type = "permute") %>% 
  calculate(stat = "diff in means", 
            order = c("2014", "2012")
            ) 

# Step 3: Calculate the p-value based on permuted differences in means 
permute %>% # using the vector of permuted differences in means
  get_p_value(obs_stat = obs_diff, # declare the observed difference in means  
              direction = "two_sided" # state the direction for the test
              )
```


__5. Do the results of the two tests differ? Why do you believe this is the case?__ 
</br>



</br>
</br>

# Your Turn: Modeling Differences in Fatalities Between Sexes  

We could investigate if the rate of deaths (per 100,000 people) in 2014 
for males and females differs. Now, instead of comparing across years, we are 
comparing across sexes within the *same* year. 


## Gathering 2014 Data 

First, we need to gather the two columns on Male and Female fatalities in 2014. 
Using the code above, write code to:  

1. Select the `Male, 2014` and `Female, 2014` columns  
2. Transform these columns from wide format to long format, specifying a new key 
that is relevant to the variables being gathered  
3. Change the values of the variable you named (in the `key` argument of 
`gather`) to take on values of "Male" and "Female" instead  

__6. Insert your code here!__ 

```{r}

```  

</br>
</br>

Next, we need to investigate what type of procedure we should use to test if there 
is a difference in fatality rates for males versus females. To do this, we need 
to create plots to determine how violate the model's assumptions are.  

__7. Create a side-by-side violin plot of the fatality rates for the two sexes.__ 

```{r}


```  


</br>
</br>

__8. Determine how violated each of the assumptions for conducting a t-test is.__ 

* __Independence:__  


* __Normality of Errors/Response:__  


* __Equal Variance:__  


